---
title: "madrid"
author: "Ermioni Athanasiadi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE}
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(data.table)
library(gtsummary)
library(mice)
library(data.table)
library(lubridate)
```


```{r}
prefix <- "madrid_"
years <- seq(2001, 2018, 1)

files <- paste0(prefix, years)



temp <- list.files(path="./VDS2526_Madrid/", pattern="\\.csv$")


stations <- read.csv("./VDS2526_Madrid/stations.csv")
```



```{r}
setwd("./VDS2526_Madrid/")

allfiles <- lapply(temp, read.csv)

allfiles <- allfiles[-19] # remove stations file

stations <- read.csv("./stations.csv")

```


```{r}
View(allfiles[[1]])
```

Transform allfiles into dataframe
```{r}
allfiles_df <- bind_rows(allfiles)
```

Merge with stations file
```{r}
df <- right_join(allfiles_df, stations, by = c("station"="id"))
```

Save all time-related information as distinct variables:
Code from kaggle: https://www.kaggle.com/code/gloriousc/madrid-s-air-quality-with-arima-forecasting/script 
```{r}
df$date <- as.POSIXct(df$date, format = "%Y-%m-%d %H:%M:%S")
df$Date <- format(df$date, "%Y-%m-%d")
df$Time <- format(df$date, "%T")
df$Date <- as.POSIXct(df$Date, format = "%Y-%m-%d")

df$day <- format(df$date, "%d")
df$month <- format(df$date, "%m")
df$year <- format(df$date, "%Y")
df$hour <- format(df$date, "%H")
```


This is time series data, and we are dealing with an extremely large dataset. We have > 20 different stations, for each station, we have data from 18 years, for each year there are > 8000 observations. 

*Brainstorm analysis ideas for this dataset*

- How does pollution vary within time of the year, time of the month, time of the day?
- How does pollution differ between more elevated and less elevated stations in Madrid?
- Do different pollutants evolve differently by station? 
- Stationarity and Structure: constant means, constant variances for which stations, for which pollutants?
- Anomaly detection: are there significant points of change? 
- what is the forecast for pollution in the next 5 years? (e.g. ARIMA)




Now let's explore the missingness of single pollutants in the dataset in more detail
```{r}
tab1 <- df %>%
    select(year, station, CO) %>%
    filter(station == "28079035") %>%
    group_by(year) %>%
    summarise(
        mean_co = mean(CO, na.rm = T),
        sd_CO = sd(CO, na.rm = T),
        length = n(),
        pmiss = sum(is.na(CO)) / n() * 100
    ) %>%
    mutate(
        across(where(is.numeric), ~ round(.x, 2))
    )
tab1
```

Let's look at how missingness in pollutants varies across stations
```{r}
df_summ <- df %>% select(2:19)

gtsummary::tbl_summary(df_summ, missing_stat = "{p_miss}", by = "station")
```



We need to obtain an overview of how missingness patterns are similar across pollutants, stations and years: 

For example, for a given station and pollutant, we have complete data, at least on the year level:

```{r}
df %>%
    select(year, station, CO) %>%
    filter(station == "28079035") %>%
    group_by(year) %>%
    ggplot(aes(y = CO, x = year)) +
    geom_boxplot()
```

```{r}
draw_boxplots <- function(df, stationX) {
    df %>%
    select(year, station, CO) %>%
    filter(station == stationX) %>%
    group_by(year) %>%
    ggplot(aes(y = CO, x = year)) +
    geom_boxplot()
}
```


Now let's look at a single pollutant (CO) but at each station separately:
```{r, echo=FALSE}
stations <- df %>%
    pull(station) %>%
    unique()

lapply(stations, function(x) draw_boxplots(df, x))

```

```{r}

# heatmap given the pollution level for each year and for each hour of the day

stat = "28079039"


heatmap_example <- df %>%
    filter(station == stat) %>%
    group_by(year, hour) %>%
    summarise(
        `Mean Pollution` = mean(O_3, na.rm = T)
    ) %>%
    ggplot(aes(x = year, y = hour, fill = `Mean Pollution`)) +
    geom_tile() +
    scale_fill_viridis(discrete=FALSE) +
    theme_ipsum()

ggsave("heatmap_example.png", heatmap_example, width = 5, height = 4)

df %>%
    filter(station == stat) %>%
    group_by(year, hour) %>%
    summarise(
        mean_pollution = mean(NO_2, na.rm = T)
    ) %>%
    ggplot(aes(x = year, y = hour, fill = mean_pollution)) +
    geom_tile()

```

```{r}
# DEBUG THIS
meta_cols <- c(
  "source_year", "source_file", "date", "station", "datetime",
  "year", "month", "month_label", "day", "hour", "wday_label",
  "season", "day_night"
)

pollutant_cols <- setdiff(names(df), meta_cols)

df_long <- reshape2::melt(
  df,
  id.vars = meta_cols[meta_cols %in% names(df)],
  measure.vars = pollutant_cols,
  variable.name = "pollutant",
  value.name = "value",
  variable.factor = FALSE
)

df_long[, pollutant := as.character(pollutant)]

missing_by_pollutant <- df_long[, .(
  n_total = .N,
  n_missing = sum(is.na(value)),
  pct_missing = 100 * mean(is.na(value))
), by = pollutant][order(-pct_missing)]
```

```{r}
pollutant_cols <- c("BEN","CO","EBE","MXY","NMHC","NO_2","NOx",
                "OXY","O_3","PM10","PXY","SO_2","TCH","TOL",
                "PM25","NO","CH4")

miss_long <- df %>%
    select(
        station, year, all_of(pollutant_cols) 
    ) %>%
    pivot_longer(all_of(pollutant_cols), names_to = "pollutant", values_to = "value") %>%
    group_by(station, year, pollutant) %>%
    summarise(
        n_total = n(),
        n_miss = sum(is.na(value)),
        pct_miss = 100*n_miss / n_total,
        .groups = "drop"
    )

miss_long %>% arrange(desc(pct_miss), station, year, pollutant) %>% View()
```

Ok, lets see which stations suffered the most missingness:

Stations with no observations at all for a given pollutant & year:
28079004
28079008
28079011
28079016
28079017
28079018
28079024


```{r}
station_summ <- miss_long %>%
  group_by(station, pollutant) %>%
  summarise(pct_miss = mean(pct_miss), .groups = "drop")

ggplot(station_summ, aes(x = pollutant, y = station, fill = pct_miss)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  scale_y_discrete(limits = rev(unique(station_summ$station))) +  # rev() puts first station at top
  scale_fill_gradientn(
    colours = c("#2166ac","#f7f7f7","#d73027"),
    values  = scales::rescale(c(0, 50, 100)),
    limits  = c(0, 100),
    name    = "% missing"
  ) +
  labs(
    title = "Missingness by station and pollutant",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid   = element_blank()
  )
```
We comment on the missingness in the data based on the pollutants:

BEN
CO
EBE
MXY
NMHC
NO2     largely observed in all stations!
NOx
OXY
O3
PM10
PXY     Only observed in 2 stations: 28079024, 28079035
SO2
TCH
TOL
PM25
NO      some missingness, but for now station missing completely
CH4     was only observed in 3 stations: 28079008, 28079024, 28079055


Lets look only at the pollutants which are listed in he WHO guidelines:
https://www.who.int/teams/environment-climate-change-and-health/air-quality-and-health/health-impacts/types-of-pollutants 

PM10
PM25
CO
NO2
SO2
O3




```{r}
year_summ <- miss_long %>%
  group_by(year, pollutant) %>%
  summarise(pct_miss = mean(pct_miss), .groups = "drop")

ggplot(year_summ, aes(x = pollutant, y = year, fill = pct_miss)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  scale_y_discrete(limits = rev(unique(year_summ$year))) +  # rev() puts first station at top
  scale_fill_gradientn(
    colours = c("#2166ac","#f7f7f7","#d73027"),
    values  = scales::rescale(c(0, 50, 100)),
    limits  = c(0, 100),
    name    = "% missing"
  ) +
  labs(
    title = "Missingness by year and pollutant",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid   = element_blank()
  )
```

We have data from all the years for NO2. Also, O3 and PM10 are largely observed in each year. After that, CO and SO2 are available.
NO only from 2011 onwards, while NOx only until 2010.

We can look at this pattern for each station separately:


```{r}
draw_heatmaps <- function(df, stationX) {
    year_summ <- miss_long %>%
    filter(station == stationX) %>%
    group_by(year, pollutant) %>%
    summarise(pct_miss = mean(pct_miss), .groups = "drop")

    ggplot(year_summ, aes(x = pollutant, y = year, fill = pct_miss)) +
      geom_tile(colour = "white", linewidth = 0.4) +
      scale_y_discrete(limits = rev(unique(year_summ$year))) +  # rev() puts first station at top
      scale_fill_gradientn(
        colours = c("#2166ac","#f7f7f7","#d73027"),
        values  = scales::rescale(c(0, 50, 100)),
        limits  = c(0, 100),
        name    = "% missing"
      ) +
      labs(
        title = paste0("Missingness by year and pollutant for station ", stationX),
        x = NULL, y = NULL
      ) +
      theme_minimal(base_size = 11) +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid   = element_blank()
      )
    
}
```


```{r, echo=FALSE}
stations <- df %>%
    pull(station) %>%
    unique()

lapply(stations, function(x) draw_heatmaps(df, x))

```

Station 28079024 has the most observations!
Also good candidates are station 28079008


AIQ indicator (check out formula)


```{r}
draw_heatmaps_pollutant <- function(df, pollutantX) {

    
    
    year_summ <- miss_long %>%
    filter(pollutant == pollutantX) %>%
    group_by(year, station) %>%
    summarise(pct_miss = mean(pct_miss), .groups = "drop")

    ggplot(year_summ, aes(x = station, y = year, fill = pct_miss)) +
      geom_tile(colour = "white", linewidth = 0.4) +
      scale_y_discrete(limits = rev(unique(year_summ$year))) +  # rev() puts first station at top
      scale_x_discrete(limits = rev(unique(year_summ$station))) +  # rev() puts first station at top
      scale_fill_gradientn(
        colours = c("#2166ac","#f7f7f7","#d73027"),
        values  = scales::rescale(c(0, 50, 100)),
        limits  = c(0, 100),
        name    = "% missing"
      ) +
      labs(
        title = paste0("Missingness by year and station for pollutant ", pollutantX),
        x = NULL, y = NULL
      ) +
      theme_minimal(base_size = 11) +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid   = element_blank()
      )
    
}
```


```{r}
who_agq_pollutants <- c("PM10", "PM25", "CO", "NO_2", "SO_2", "O_3")
lapply(who_agq_pollutants, function(x) draw_heatmaps_pollutant(df, x))
```

From these data, it seems reasonable that all conclusions about air pollution will have to rely on a different subset of stations for each pollutant. This is suboptimal, as it limits comparability across pollutants. It should be made clear in the visualisations, which stations were included in the analysis of each pollutant.

NO2 is the only pollutant that has been fully observed.

Alternatively, we could reduce our analysis on a subset of up to e.g. 5 stations for which we have observed data for all pollutants (except PM25, which is largely unobserved).

station 28079024 is the best station in terms of completeness. It even carries data for PM25 at least from year 2003 onwards. 

The following stations have data for the remaining 5 pollutants:
28079008


Let's prepare the dataset that will be used for the subsequent visualisations:

```{r}
df_long <- df %>%
    pivot_longer(all_of(pollutant_cols), names_to = "pollutant", values_to = "value") 

```

```{r}
df_long_WHOpollutants <- df_long %>%
    dplyr::filter(pollutant %in% who_agq_pollutants)
```


Now lets calculate the AQI index based on the WHO guideline:

https://document.airnow.gov/technical-assistance-document-for-the-reporting-of-daily-air-quailty.pdf 

```{r}
get_aqi <- function(C, BP_hi, BP_lo, I_hi, I_lo)  {

    pollutant_value  <- (I_hi - I_lo) / (BP_hi - BP_lo) * (C - BP_lo) + I_lo 
    return(pollutant_value)
}

```


This is the easiest way to summarise the data based on the AQI procedure, but it
is not feasible for the size of this dataset. We will use data.table which works faster (see below)
```{r eval=FALSE, include=FALSE}
# add the AQI
df_long_daily <- df_long_WHOpollutants %>%
  select(station, Date, hour, pollutant, value) %>%
  group_by(station, Date, pollutant) %>%
  summarise(
    conc = case_when(
      pollutant == "PM10"  ~ mean(value, na.rm=TRUE),   # 24h mean
      pollutant == "PM2.5" ~ mean(value, na.rm=TRUE),   # 24h mean
      pollutant == "NO_2"  ~ max(value, na.rm=TRUE),    # 1h max
      pollutant == "SO_2"  ~ max(value, na.rm=TRUE),    # 1h max
      pollutant == "CO"    ~ max(zoo::rollmean(value, 8, fill=NA, align="right"), na.rm=TRUE),  # 8h rolling max
      pollutant == "O_3"   ~ max(zoo::rollmean(value, 8, fill=NA, align="right"), na.rm=TRUE)  # 8h rolling max
    )
  ) %>%
  # Then apply get_aqi() with the right breakpoints per pollutant
  mutate(aqi_sub = get_aqi(conc, BP_hi, BP_lo, I_hi, I_lo))
```


```{r}
dt <- as.data.table(df_long_WHOpollutants)

# Subgroup 1: mean (PM10, PM2.5)
dt_mean <- dt[pollutant %in% c("PM10", "PM2.5"),
              .(conc = mean(value)),
              by = .(station, Date, pollutant)]

# Subgroup 2: 1h max (NO2, SO2)
dt_max <- dt[pollutant %in% c("NO_2", "SO_2"),
             .(conc = max(value)),
             by = .(station, Date, pollutant)]

# Subgroup 3: 8h rolling mean then daily max (O3, CO)
dt_roll <- dt[pollutant %in% c("O_3", "CO")]
dt_roll[, roll8 := frollmean(value, 8, fill = NA, align = "right"),
        by = .(station, pollutant)]  # first roll across full time series per station
dt_roll <- dt_roll[, .(conc = max(roll8, na.rm = TRUE)), # then get max per day
                   by = .(station, Date, pollutant)]

# combine subgroups
dt_daily <- rbindlist(list(dt_mean, dt_max, dt_roll))
```


We will recreate the Date variables now:
```{r}
dt_daily$day <- format(dt_daily$Date, "%d")
dt_daily$month <- format(dt_daily$Date, "%m")
dt_daily$year <- format(dt_daily$Date, "%Y")
dt_daily$hour <- format(dt_daily$Date, "%H")
```

Now we are ready to calculate the AQI on our daily-dataset:

The tricky part is that we need to compare each pollution value with an interval. The trick is to turn the pollution value into an interval itself and then compare two intervals, referred to as "overlap joins". Two intervals overlap if the following condition holds:

start_A <= end_B && start_B <= end_A

To start we have to create lookup tables to look up the breakpoints for each pollutant:
```{r}

bp_o3 <- data.table(  #based on 8-hour interval (1h also possible)
  pollutant = "O_3",
  c_low  = c(0.000, 0.055, 0.071, 0.086, 0.106, 0.201),
  c_high = c(0.054, 0.070, 0.085, 0.105, 0.200, 604),
  i_low  = c(0, 51,101,151,201, 301),
  i_high = c(50,100,150,200,300, 400)
)

bp_pm10 <- data.table(
  pollutant = "PM10",
  c_low  = c(0, 55, 155, 255, 355, 425, 505),
  c_high = c(54,154,254,354,424,504,604),
  i_low  = c(0, 51,101,151,201,301,401),
  i_high = c(50,100,150,200,300,400,500)
)

bp_pm25 <- data.table(
  pollutant = "PM2.5",
  c_low  = c(0.0, 12.1, 35.5, 55.5, 150.5, 250.5, 350.5),
  c_high = c(12.0, 35.4, 55.4, 150.4, 250.4, 350.4, 500.4),
  i_low  = c(0, 51, 101, 151, 201, 301, 401),
  i_high = c(50,100,150,200,300,400,500)
)

bp_co <- data.table(
  pollutant = "CO",
  c_low  = c(0.0, 4.5, 9.5, 12.5, 15.5, 30.5, 40.5),
  c_high = c(4.4, 9.4, 12.4, 15.4, 30.4, 40.4, 50.4),
  i_low  = c(0, 51,101,151,201,301,401),
  i_high = c(50,100,150,200,300,400,500)
)

bp_so2 <- data.table(
  pollutant = "SO_2",
  c_low  = c(0, 36, 76, 186, 305, 605, 805),
  c_high = c(35, 75, 185, 304, 604, 804, 1004),
  i_low  = c(0, 51,101,151,201,301,401),
  i_high = c(50,100,150,200,300,400,500)
)

bp_no2 <- data.table(
  pollutant = "NO_2",
  c_low  = c(0, 54, 101, 361, 650, 1250, 1650),
  c_high = c(53,100,360,649,1249,1649,2049),
  i_low  = c(0, 51,101,151,201,301,401),
  i_high = c(50,100,150,200,300,400,500)
)

bp_all <- rbindlist(list(
  bp_pm10,
  bp_pm25,
  bp_o3,
  bp_co,
  bp_so2,
  bp_no2
))

rm(bp_pm10, bp_pm25, bp_o3, bp_co, bp_so2, bp_no2)
```


```{r}
setDT(dt_daily)

dt_daily[, conc_num := as.numeric(conc)]

dt_aqi <- dt_daily
dt_aqi <- dt_daily[!is.na(conc_num)]  # deletes ~ 170000 obs. check if all PM2.5 values are excluded in this step

dt_aqi[, `:=`(
  conc_low = conc_num,
  conc_high = conc_num
)]

setkey(dt_aqi, pollutant, conc_low, conc_high)
setkey(bp_all, pollutant, c_low, c_high)

aqi_joined <- foverlaps(
  dt_aqi,
  bp_all,
  by.x = c("pollutant", "conc_low", "conc_high"),
  by.y = c("pollutant", "c_low", "c_high"),
  nomatch = NA
)

aqi_joined[, AQI :=
  round(
    ((i_high - i_low) / (c_high - c_low)) *
      (conc_num - c_low) + i_low
  )
]
```
We can now also determine the overall AQI and summarise over all pollutants. Lets also save which pollutant is responsible
```{r}
daily_overall <- aqi_joined[
  ,
  .(
    overall_AQI = max(AQI, na.rm = TRUE),
    responsible_pollutant = pollutant[which.max(AQI)]
  ),
  by = .(station, Date)
]
```

Now lets create a new category for visualising later:
```{r eval=FALSE, include=FALSE}

### 
aqi_joined[,
  
  `:=`(

    AQI_color = fifelse(
      AQI <= 50, "#00E400",
      fifelse(
        AQI <= 100, "#FFFF00",
        fifelse(
          AQI <= 150, "#FF7E00",
          fifelse(
            AQI <= 200, "#FF0000",
            fifelse(
              AQI <= 300, "#8F3F97",
              "#7E0023"
            ))))))
  ]
```

```{r}
get_aqi_categorical <- function(aqi_var) {

  cut(
    aqi_var,
    breaks = c(-Inf, 50, 100, 150, 200, 300, Inf),
    labels = c(
      "Good",
      "Moderate",
      "Unhealthy for Sensitive Groups",
      "Unhealthy",
      "Very Unhealthy",
      "Hazardous"
    )
  )
}
```

```{r}
aqi_joined[, AQI_categorical := get_aqi_categorical(AQI)]

daily_overall[, AQI_categorical := get_aqi_categorical(overall_AQI)]
```

Let's save our dataset to import it later for the visualisations:

```{r}

```


So now we are ready to get some visualisations:
```{r}
aqi_cols <- c(
  "Good" = "#00E400",
  "Moderate" = "#FFFF00",
  "Unhealthy for Sensitive Groups" = "#FF7E00",
  "Unhealthy" = "#FF0000",
  "Very Unhealthy" = "#8F3F97",
  "Hazardous" = "#7E0023"
)


ggplot(
  daily_overall,
  aes(Date, overall_AQI, color = AQI_categorical)
) +
  geom_point(alpha = 0.7) +
  scale_color_manual(values = aqi_cols) +
  theme_minimal() +
  labs(
    title = "Daily AQI",
    y = "AQI",
    color = "Category"
  )

```

```{r}
ggplot(
  daily_overall,
  aes(Date, factor(station), fill = AQI_categorical)
) +
  geom_tile() +
  scale_fill_manual(values = aqi_cols) +
  theme_minimal() +
  labs(
    x = "Date",
    y = "Station",
    fill = "AQI"
  )
```
Ok, from further below we know that this is largely driven by Ozone levels.
Let's look at the pollutant-level-AQIs:
```{r}
aqi_pollutants <- c("PM10", "CO", "NO_2", "SO_2", "O_3")


draw_aqi_heatmaps_pollutant <- function(df, pollutantX) {
    aqi_joined %>% 
        filter(pollutant == pollutantX) %>%
        nrow(.) %>% print()
    aqi_joined %>% 
        filter(pollutant == pollutantX) %>%
    ggplot(
      aes(Date, factor(station), fill = AQI_categorical)
    ) +
      geom_tile() +
      scale_fill_manual(values = aqi_cols) +
      theme_minimal() +
      labs(
        title = paste0("AQI value for pollutant ", pollutantX),
        x = "Date",
        y = "Station",
        fill = "AQI"
      ) 
}

lapply(aqi_pollutants, function(x) draw_aqi_heatmaps_pollutant(df, x))
```

Same information, but more densely presented:
```{r}

ggplot(
  aqi_joined,
  aes(Date, factor(station), fill = AQI)
) +
  geom_tile() +
  facet_wrap(~ pollutant) +
  scale_fill_gradientn(colors = unname(aqi_cols)) +
  theme_minimal()

ggplot(
  aqi_joined,
  aes(Date, factor(station), fill = AQI_categorical)
) +
  geom_tile() +
  facet_wrap(~ pollutant) +
  scale_fill_manual(values = aqi_cols) +
  theme_minimal()


```


There is no apparent strong heterogeneous change of AQI over time depending on the station. Thus, we will not look at the data on the station level anymore for the subsequent analyses. We are interested in the overall air quality levels in Madrid. 

In addition, we know that through selective missingness patterns, each station is composed of a different number of observations and pollutants. Thus, comparing overall AQI levels on the station level seems, to say the least, questionable.

```{r}
ggplot(
  aqi_joined,
  aes(Date, AQI, color = pollutant)
) +
  geom_line(alpha = 0.4) +
  facet_wrap(~ pollutant, scales = "free_y") +
  theme_minimal()
```


```{r}

aqi_joined[, `:=`(
  year = lubridate::year(Date),
  month = lubridate::month(Date, label = T),
  day = lubridate::day(Date)
)]
```

```{r}
ggplot(
  aqi_joined[pollutant == "PM10"],
  aes(day, month, fill = AQI)
) +
  geom_tile() +
  facet_wrap(~ year) +
  scale_fill_gradientn(colors = unname(aqi_cols))

ggplot(
  aqi_joined[pollutant == "NO_2"],
  aes(day, month, fill = AQI)
) +
  geom_tile() +
  facet_wrap(~ year) +
  scale_fill_gradientn(colors = unname(aqi_cols))

ggplot(
  aqi_joined[pollutant == "CO"],
  aes(day, month, fill = AQI)
) +
  geom_tile() +
  facet_wrap(~ year) +
  scale_fill_gradientn(colors = unname(aqi_cols))

ggplot(
  aqi_joined[pollutant == "O_3"],
  aes(day, month, fill = AQI)
) +
  geom_tile() +
  facet_wrap(~ year) +
  scale_fill_gradientn(colors = unname(aqi_cols))

ggplot(
  aqi_joined[pollutant == "SO_2"],
  aes(day, month, fill = AQI)
) +
  geom_tile() +
  facet_wrap(~ year) +
  scale_fill_gradientn(colors = unname(aqi_cols))


```

Overall this is good news: Air quality has improved over the years based on these 5 pollutants. Except for Ozone, which is at a bad level throughout the years. 

We could design this as interactive plot. The interactive element should be the Pollutant, chosen by the user as dropdown. The evolution over the years should always appear within one given visualisation. 

First, we need to adapt the visualisation to make it more readable:

```{r}

# remove years with NA -> this should solve months NAs too


# resort the months y-axis: from jan-dez


# more sparse data presentation: only present years 2001, 2005, 2009, 2013, 2017

# or even top 3: 2001, 2009, 2017


# 



```




```{r}
aqi_joined[, season := fifelse(
  month(Date) %in% c(12,1,2), "Winter",
  fifelse(
    month(Date) %in% c(3,4,5), "Spring",
    fifelse(
      month(Date) %in% c(6,7,8), "Summer",
      "Autumn"
    )
  )
)]

ggplot(
  aqi_joined,
  aes(season, AQI, fill = pollutant)
) +
  geom_boxplot(outlier.alpha = 0.1)
```


```{r}

ggplot(
  daily_overall,
  aes(Date, factor(station), fill = responsible_pollutant)
) +
  geom_tile()
```

This is an important insight. It shows us that Overall Pollution Level as measured by the Overall AQI is almost entirely driven by Ozone levels, and, partly, by NO2.

We should get the same result if we use the aqi_joined dataset and summarise over pollutant, via choosing the maximum.




Now lets get an idea of our summarised dataset:
```{r}
dt_daily %>%
    group_by(station) %>%
    summarise(
        n = n()
    )
dt_daily %>%
    group_by(pollutant) %>%
    summarise(
        n = n()
    )

dt_daily %>%
    group_by(Date) %>%
    summarise(
        n = n()
    )

dt_daily %>%
    group_by(station, pollutant) %>%
    summarise(
        n = n()
    )

dt_daily %>%
    group_by(year) %>%
    summarise(
        n = n()
    )

dt_daily %>%
    group_by(year, station, pollutant) %>%
    summarise(
        n = n()
    )
```



```{r}
saveRDS(df_long_WHOpollutants, file = "./../data/df_long_WHOpollutants.RDS")
saveRDS(dt_daily, file = "./../data/dt_daily.RDS")
saveRDS(aqi_joined, file = "./../data/aqi_joined.RDS")
saveRDS(daily_overall, file = "./../data/daily_overall.RDS")



```


*ALTERNATIVE APPROACH*
The AirMonitor package offers functionality to calculate AQI directly - however, this is available for PM2.5 and O_3.

```{r}
library(AirMonitor)


dt_daily_wide <- dcast(dt_daily, station + Date ~ pollutant, value.var = "conc")

```



```{r}
lapply(who_agq_pollutants, function(x) draw_heatmaps_pollutant(df_long_WHOpollutants, x))
```


